---
title: "Task 4 Analysis"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
candy_cleaned <- read_csv(here("clean_data/boing_boing_cleaned.csv"))
```


1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
# First we select all the columns with ratings; this is what we want to count.
candy_cleaned_ratings_only <- candy_cleaned %>% 
  select(x100_grand_bar:sandwich_sized_bags_filled_with_boo_berry_crunch)

# Then we sum all the Not NA values, giving us the total number of candy ratings!
sum(!is.na(candy_cleaned_ratings_only))

```

2. What was the average age of people who are going out trick or treating?

```{r}
# We filter so we only get those who went out trick or treating, then took the mean of the age column; removing NAs so we could perform this calculation!

candy_cleaned %>% 
  filter(trick_or_treating == "Yes") %>% 
  filter(age < 120) %>% 
  summarise(avg_age_of_trick_or_treaters = mean(age, na.rm = TRUE))
```





3. What was the average age of people who are not going trick or treating?

```{r}
candy_cleaned %>% 
  filter(trick_or_treating == "No") %>% 
  filter(age < 120) %>% 
  summarise(avg_age_of_not_trick_or_treaters = mean(age, na.rm = TRUE))
```



4. For each of joy, despair and meh, which candy bar received the most of these ratings?

# The issue here is we're working in both columns and rows, it's tricky. I have counted the amount of JOY per column but I can't map that to the corresponding candy column. I suspect some sort of function could be  written to increment a count of each JOY for each column, though frankly I've no idea how to do that.

Alternatively, we could create some sort of index for the candy columns which would map to our total_joy_per_candy vector, then search for the highest no. of ratings in that vector - we could then easily  match them. 

```{r}

# JOY

total_joy_per_candy <- c(str_count(candy_cleaned_ratings_only, "JOY"))

index_joy <- which(total_joy_per_candy == max(total_joy_per_candy))

candy_with_most_joy <- candy_cleaned_ratings_only[index_joy]


# MEH

total_meh_per_candy <- c(str_count(candy_cleaned_ratings_only, "MEH"))

index_meh <- which(total_meh_per_candy == max(total_meh_per_candy))

candy_with_most_meh <- candy_cleaned_ratings_only[index_meh]


# DESPAIR

total_despair_per_candy <- c(str_count(candy_cleaned_ratings_only, "DESPAIR"))

index_despair <- which(total_despair_per_candy == max(total_despair_per_candy))

candy_with_most_despair <- candy_cleaned_ratings_only[index_despair]


colnames(candy_with_most_joy)
colnames(candy_with_most_meh)
colnames(candy_with_most_despair)



```



5. How many people rated Starburst as despair?

```{r}
candy_cleaned %>% 
  filter(starburst == "DESPAIR") %>% 
  nrow()
```

For the next three questions, count despair as -1, joy as +1, and meh as 0.

How do we convert JOY to 1, despair to -1 etc? Probably want to use str_detect and a function, that replaces string with a number when the string match is TRUE.

Prequel to next 3 questions:

```{r}


# candy_cleaned %>% 
#     mutate(across(matches('[A-Za-z]'), ~ case_when(. == "DESPAIR" ~ -1, 
#                      . == 'MEH' ~ 0,
#                      . == "JOY" ~ 1,
#                      TRUE ~ NA_real_)))
# 
# 
# 
# candy_cleaned %>% 
#   mutate(across(.cols = everything(), .fns = NULL, str_replace_all(candy_cleaned, "DESPAIR", "-1")))
# 
# 
# while (candy_cleaned$x100_grand_bar) == "DESPAIR"){
#   candy_cleaned$x100_grand_bar <- "-1"
# }


candy_cleaned <- candy_cleaned %>% 
  mutate(x100_grand_bar = gsub("DESPAIR", "-1", candy_cleaned$x100_grand_bar))

candy_cleaned <- candy_cleaned %>% 
  mutate(x100_grand_bar = gsub("MEH", "0", candy_cleaned$x100_grand_bar))

candy_cleaned <- candy_cleaned %>% 
  mutate(x100_grand_bar = gsub("JOY", "1", candy_cleaned$x100_grand_bar))

candy_cleaned <- candy_cleaned %>% 
  mutate(x100_grand_bar = as.numeric(x100_grand_bar))





candy_cleaned <- candy_cleaned %>%
  mutate(across(x100_grand_bar:sandwich_sized_bags_filled_with_boo_berry_crunch, gsub("JOY", "1", candy_cleaned)))






```




6. What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}

```


7. What was the most popular candy bar in each year?

```{r}

```


What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}

```

