---
title: "Task 1"
output: html_notebook
---

Here we load in tidyverse and the here package - allowing our code to be reproducible. Others using this project should have no issue loading in the cleaned data.
```{r}
library(tidyverse)
library(here)
decathlon <- read_csv(here("clean_data/decathlon_cleaned.csv"))
```

Analysis Questions:

1. Who had the longest long jump seen in the data?
```{r}
decathlon %>% 
  select(long_jump, names) %>% 
  filter(long_jump == max(long_jump))

# CLAY has the longest long jump.
```



2. What was the average 100m time in each competition?
```{r}
decathlon %>% 
  group_by(competition) %>% 
  summarise(average_100m_time = round(mean(x100m), 2))

# The average 100m time is 11.18 seconds for Decastar, and 10.92 for OlympicG.
```


3. Who had the highest total points across both competitions?
```{r}
s
```



4. What was the shot-put scores for the top three competitors in each competition?
```{r}

```



5. What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}

```

